s---
title: "Cleaning"
output: pdf_document
authors: Adam, Minyue, Nael, and Sonja
date: 11-Feb-2020
---

Load the combined covariate, metabolite, and worklist data.
```{r}
setwd("~/dissertation/dissertation/mtb")
source_rmd <- function(f){library(knitr); source(purl(f, tempfile(), quiet=TRUE))}
source_rmd("load.Rmd")
```

Load impute package needed for "impute.knn".
```{r include=FALSE}
install_load("BiocManager")
if (!require("impute")) BiocManager::install("impute")
library("impute")
```

Clean our data frame by removing duplicate metabolites (features) and samples by only keeping the most recent. Also remove any blank samples that were used for quality control.
```{r}
data_clean <- data %>%
  select(-`POS_855.8915@6.0343328:2`) %>% # Remove duplicate metabolite (feature)
  rename(`POS_855.8915@6.0343328`=`POS_855.8915@6.0343328:1`) %>% #renames to remove colon
  group_by(Indiv.ID) %>% # Remove duplicate samples
  summarise_all(last) %>% # keep last individual
  filter(!is.na(case)) # Remove blank samples
```

Extract metabolite (feature) information from cleaned table. Remove any features with more than 50% missing data or zero variance. Finally, log transform the metabolite (feature) data. 
```{r}
data_metabolite <- data_clean %>%
  select(matches("POS_.*|NEG_.*")) %>% # Select anything that begins with "POS" or "NEG"
  select_if(~mean(is.na(.)) <= 0.50) %>% # Remove features with more than 50% NA's; dot means anything that's being pushed into it
  select_if(~length(unique(.)) > 1) %>% # Remove all metabolites that have zero variance
  mutate_all(log10) # Convert all metabolites (features) to log scale
```

Remove any participants (rows) with more than 80% missing data.
```{r}
data_clean <- data_clean[which(rowMeans(is.na(data_metabolite)) < 0.20),]
data_metabolite <- data_metabolite[which(rowMeans(is.na(data_metabolite)) < 0.20),]
```

Scale the metabolite data to the unit scale allowing K nearest neighbours to correctly idenitfy neighbours.
```{r}
data_metabolite <- data_metabolite %>%
  mutate_all(scale) # scaling is not uniform
```

Impute missing metabolite data using K nearest neighbours.
```{r}
data_metabolite.imputed <- data_metabolite %>%
  as.matrix() %>% # because impute.knn only works on matrices
  impute.knn() %>% 
  .$data %>% # accesses the data
  as.data.frame() # pushes the data into a dataframe
```

Unscale the metabolite back to the log scale.
```{r}
data_metabolite.imputed <- data_metabolite.imputed %>%
  map2_df(., data_metabolite,
          function(x, y) x * attr(y, "scaled:scale") + attr(y, "scaled:center")) 
# map2df takes two dataframes and applies a function to each unit in the dataframe
#function x is the dot, function y is the data_metabolite
```

Replace the metabolite data in our cleaned data frame with our freshly imputed data.
```{r}
data_clean <- data_clean %>%
  select(-matches("POS_.*|NEG_.*")) %>% # select everything that isn't a metabolite
  cbind(data_metabolite.imputed)
```

Remove our processing variables to avoid polluting the global environment.
```{r}
rm(data_metabolite, data_metabolite.imputed)
```
