---
title: "Denoising"
output: pdf_document
authors: Adam, Minyue, Nael, and Sonja
date: 19-Feb-2020
---

Load the cleaned covariate, metabolite, and worklist data.
```{r}
source_rmd <- function(f){library(knitr); source(purl(f, tempfile(), quiet=TRUE))}
source_rmd("clean.Rmd")
```

Load mixOmics package to calculate mixed linear models.
```{r}
install_load("omics")
```

Filter only females and split data into positive and negative metabolites.
```{r}
data_clean_female <- data_clean %>% filter(gender == "Female")

data_clean_neg <- data_clean_female %>% select(matches("NEG_.*"))
data_clean_pos <- data_clean_female %>% select(matches("POS_.*"))
```


```{r}
data_mlmer <- data_clean_female %>% 
  mutate(Position.pos=factor(Position.pos), 
         Position.neg=factor(Position.neg), 
         Plate.pos=factor(Plate.pos),
         Plate.neg=factor(Plate.neg))
```

Use mixed linear models to denoise positive metabolite data on plate, position, age, and BMI. For cases missing BMI data we denoise on the other variables - as there is no obvious way of manually calculating residuals from mlmer we use the non ideal solution of calling mlmer again and combining the results.


```{r}
cov1 = data_mlmer[,c(1:35)]
```


```{r}
denoised_pos = NULL

#only female! no need to adjust gender
for (k in c(1:ncol(data_clean_pos))) {
  model = lmer(data_clean_pos[,k] ~ age.sample + bmi + centre + education_plco  +  smok_duration + smok_intensity + smoking_status + (1 | Plate.pos) + (1|Position.pos) + Monocytes + B + CD4T + CD8T +NK + case, 
             data = cov1, 
             REML = FALSE,
             control = lmerControl(check.conv.singular = .makeCC(action ="ignore",tol = 1e-04)))
  
  
  
  model0 = lmer(data_clean_pos[,k] ~ age.sample  + bmi + centre + education_plco  +  smok_duration + smok_intensity + smoking_status + (1 | Plate.pos) + (1|Position.pos) + Monocytes + B + CD4T + CD8T +NK, data = cov1, REML = FALSE,control = lmerControl(check.conv.singular = .makeCC(action = "ignore", tol = 1e-04)))
  
  
  
  pvalue_pooled = c(pvalue_pooled, anova(model, model0)$"Pr(>Chisq)"[2]) 
  
  beta = fixef(model)[c("(Intercept)", "case")]

  Beta_pooled = c(Beta_pooled, fixef(model)["type1"])
  
 X = cbind(rep(1,length(cov1$case)),
           as.numeric(cov1$case))

denoised_pos = cbind(denoised_pos,(X %*% beta + resid(model)))
}
```

```{r}
#adc1 <- adc[,-1]

colnames(denoised_pos) = colnames(data_clean_pos) 
rownames(denoised_pos) = rownames(data_clean_pos)

denoised_pos1 = as.data.frame(denoised_pos)

setwd("~/dissertation/dissertation/mtb")
write.csv(denoised_pos1,"mtb_pos_denoised.csv")
```




```{r}
denoised_neg = NULL

for (k in c(1:ncol(data_clean_neg))) {
  model = lmer(data_clean_neg[,k] ~ age.sample + bmi + centre + education_plco  +  smok_duration + smok_intensity + smoking_status + (1 | Plate.neg) + (1|Position.neg) +  Monocytes + B + CD4T + CD8T + NK + case, 
             data = cov1, 
             REML = FALSE,
             control = lmerControl(check.conv.singular = .makeCC(action ="ignore",tol = 1e-04)))
  
  
  
  model0 = lmer(data_clean_neg[,k] ~ age.sample  + bmi + centre + education_plco  +  smok_duration + smok_intensity + smoking_status + (1 | Plate.neg) + (1|Position.neg) + Monocytes + B + CD4T + CD8T + NK, data = cov1, REML = FALSE,control = lmerControl(check.conv.singular = .makeCC(action = "ignore", tol = 1e-04)))
  
  
  
  pvalue_pooled = c(pvalue_pooled, anova(model, model0)$"Pr(>Chisq)"[2]) 
  
  beta = fixef(model)[c("(Intercept)", "case")]

  Beta_pooled = c(Beta_pooled, fixef(model)["type1"])
  
 X = cbind(rep(1,length(cov1$case)),
           as.numeric(cov1$case))

denoised_neg = cbind(denoised_neg,(X %*% beta + resid(model)))
}

colnames(denoised_neg) = colnames(data_clean_neg) 
rownames(denoised_neg) = rownames(data_clean_neg)

denoised_neg1 = as.data.frame(denoised_neg)

setwd("~/dissertation/dissertation/mtb")
write.csv(denoised_neg1,"mtb_neg_denoised.csv")
```
