```{r}
install_load <- function(...) {
  for (package in list(...)) {
    require(package, character.only=TRUE) || install.packages(package)
    library(package, character.only=TRUE)
  }
}
```

Load the packages required for reading and manipulating the data.
```{r}
install_load("tidyverse", "readxl")
```

Load covariate, metabolite, and worklist data into relevant variables.
```{r}
setwd("~/dissertation/dissertation/mtb")
data_covariate <- read.csv("Covariates1.csv")
data_metabolite_pos <- read.csv("mtb_pos.csv")
data_metabolite_neg <- read.csv("mtb_neg.csv")
data_metabolite_worklist_pos <- read_excel("LC-MS_worklist_LunCan_RPpos.xlsx")
data_metabolite_worklist_neg <- read_excel("LC-MS_worklist_LunCan_RPneg.xlsx")
```

Combine all data into a single table named `data` to facilitate further analysis. Each row will represent a case  (including the blank samples from the metabolite data) and the columns will contain covariate, metabolite, and worklist data. Metabolite columns take the form `[Positive or Negative mode]_[Mass/Charge ratio]@[Retention time]`. Worklist columns are suffixed by either `.pos` or `.neg` to identify which mode of mass spectroscopy they relate to.
```{r}
process_metabolite <- function(df, prefix="") {
  df <- df %>% 
    column_to_rownames("Compound") %>% 
    t() %>% 
    as.data.frame() %>% 
    rename_all(function(x) paste0(prefix, x)) %>%
    rownames_to_column("Compound") %>%
    separate(Compound, c("Sample.ID", "Indiv.ID"), sep="_", extra="drop")
  return(df)
}

data <- data_covariate %>%
  mutate(Indiv.ID=as.character(Indiv.ID)) %>%
  full_join(process_metabolite(data_metabolite_pos, "POS_"), by="Indiv.ID") %>%
  full_join(process_metabolite(data_metabolite_neg, "NEG_"), by=c("Indiv.ID", "Sample.ID")) %>%
  left_join(data_metabolite_worklist_pos, by=c("Indiv.ID"="Sample Name")) %>%
  left_join(data_metabolite_worklist_neg, by=c("Indiv.ID"="Sample Name"), suffix=c(".pos", ".neg"))
```
